Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-14 SI.tex
cover letter and citations
Final edits before sending to Andrei

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by an MSM constructed at the $\tau$ = 1.2 ns lag time (dashed blue line) versus those computed from an MSM constructed at the particular lag time (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

The NN model assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions,
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}.
\end{equation}
$\Delta H_{NN}^{\circ}$ and $\Delta S_{NN}^{\circ}$ are assumed to be independent of temperature $T$, and are computed from the sum over all NN terms in a given configuration,
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ},
\end{equation}
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}.
\end{equation}                          
$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA, $\Delta H_{DE, j}^{\circ}$ corresponds to the enthalpic contribution from a given dangling end (DE) base next to the duplex\citep{Santalucia2004TM}, and $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. A schematic illustration of the NN and DE contributions for a fully hybridized duplex and an out-of-register shifted state is presented in Fig.~\ref{fig:SIFig4}. $\Delta S_{NN}^{\circ}$ is constructed in an analogous manner from NN, dangling end, and initiation terms, plus a symmetry term, $\Delta S_{sym}^{\circ}$, that applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry.

@@

\bibliography{references_210505}
%\bibliography{refs_mendeley2}

>>>>

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by $k$ applications of an MSM constructed at the $\tau$ = 1.2 ns lag time $\mathbf{P}^k(\tau)$ (dashed blue line) versus those computed from an MSM constructed at that particular lag time $\mathbf{P}(k \tau)$ (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

The NN model assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions,
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}.
\end{equation}
$\Delta H_{NN}^{\circ}$ and $\Delta S_{NN}^{\circ}$ are assumed to be independent of temperature $T$, and are computed from the sum over all NN terms in a given configuration,
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ},
\end{equation}
\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}.
\end{equation}                          
$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA, $\Delta H_{DE, j}^{\circ}$ corresponds to the enthalpic contribution from a given dangling end (DE) base next to the duplex\citep{Santalucia2004TM}, and $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. A schematic illustration of the NN and DE contributions for a fully hybridized duplex and an out-of-register shifted state is presented in Fig.~\ref{fig:SIFig6}. $\Delta S_{NN}^{\circ}$ is constructed in an analogous manner from NN, dangling end, and initiation terms, plus a symmetry term, $\Delta S_{sym}^{\circ}$, that applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry.

@@

\bibliography{references_210505}

### 2021-05-14 cover_letter/letterhead.tex
Final edits before sending to Andrei

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
%\hspace*{\longindentation} % Move the signature right
%\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.0in
\includegraphics[width=0.25\textwidth]{signature_720}
%\vskip 0.65in % Whitespace between the signature text and your name
\\
\noindent
\fromsig
\vskip -0.05in
\titleone\\
\department\\
\institution\\
\vskip -0.1in
\noindent cc: M.~Jones, B.~Ashwood, A.~Tokmakoff
}

@@

We are pleased to submit a manuscript for consideration for publication in the \textit{The Journal of the American Chemical Society} an original research \textbf{Article} entitled \textbf{``Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy''} by Michael S. Jones (U.~Chicago), Brennan Ashwood(U.~Chicago), Andrei Tokmakoff(U.~Chicago), and Andrew L.\ Ferguson (U.~Chicago). This article has not been published previously and is not under consideration in any other journal. All authors have seen and approved the submission of the manuscript.

% background and why it is important to do this.
\textbf{Background.}  A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. Although previous experimental and computational studies have identified deviations from an "all-or-nothing" hybridization/dehybridization model, few studies have integrated these approaches in order to evaluate and quantify sequence-dependent dynamics.

% what we did in this paper
\textbf{Key Results.}  In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning of the slow dynamical pathways, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. We conducted more than 1 ms of unbiased coarse-grained molecular dynamics simulations and employed deep learning techniques to construct high-resolution Markov state models as predictive and interpretable models of the sequence dependent dynamics. Our results reveal that the specific placement of interrupting G:C pairs within an otherwise repetitive AT sequence can have a profound impact on the kinetic pathways and mechanisms for association and dissociation of the DNA duplex.

% impact and importance of this paper
\textbf{Significance.} Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

% fit for the journal
\textbf{Scope.}  This work fits within the scope of \textit{The Journal of the American Chemical Society} in that it reports new fundamental physical chemical understanding of the molecular pathways and mechanisms of DNA hybridization as a function of sequence. These sequence-dependent mechanisms are of increasing importance to both fundamental biophysical and biochemical processes and rapidly developing nanotechnology applications.

@@

\textbf{\textit{Simulation.}} \noteb{Notes are just to show connection}

@@

$\circ$ \textit{Expert in multi-scale biomolecular simulation with a focus in materials science and biology. \noteb{3spn2}}

Jonathon K. Whitmer \hfill Chemical and Biomolecular Engineering, Notre Dame \\
(574) 631-1417 \hfill \url{jwhitme1@nd.edu}

$\circ$ \textit{Expert in simulations of polyelectrolytes and self-assembling systems.  \noteb{3spn2}}

@@

$\circ$ \textit{Expert in theory and simulation of multi-molecular systems.  \noteb{3spn.0 and .1}}

@@

$\circ$ \textit{Expert in simulations of DNA-protein systems \noteb{3spn2 and  CG protein modeling}}

@@

$\circ$ \textit{Expert in simulations of biochemical systems for nanotechnology applications \noteb{oxDNA}}

@@

$\circ$ \textit{Expert in molecular dynamics simulations of nucleic acids \noteb{all-atom DNA model}}

>>>>

\long\def\closing#1{
\vspace{0.1in} % Some whitespace after the letter content and before the signature
\noindent % Stop paragraph indentation
%\hspace*{\longindentation} % Move the signature right
%\parbox{\indentedwidth}{\raggedright
#1 % Print the signature text
\vskip 0.0in
\includegraphics[width=0.25\textwidth]{signature_720}
%\vskip 0.65in % Whitespace between the signature text and your name
\\
\noindent
\fromsig
\vskip -0.05in
\titleone\\
\department\\
\institution\\
\vskip -0.1in
\noindent cc: M.S.~Jones, B.~Ashwood, A.~Tokmakoff
}

@@

We are pleased to submit a manuscript for consideration for publication in the \textit{The Journal of the American Chemical Society} an original research \textbf{Article} entitled \textbf{``Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy''} by Michael S. Jones, Brennan Ashwood, Andrei Tokmakoff, and Andrew L.\ Ferguson, all from the University of Chicago. This article has not been published previously and is not under consideration in any other journal. All authors have seen and approved the submission of the manuscript.

% background and why it is important to do this.
\textbf{Background.}  A robust understanding of the sequence-dependent thermodynamics of DNA hybridization has enabled rapid advances in DNA nanotechnology. A fundamental understanding of the sequence-dependent kinetics and mechanisms of hybridization and dehybridization remains comparatively underdeveloped. Although previous experimental and computational studies have identified deviations from an ``all-or-nothing'' hybridization/dehybridization model, few studies have integrated these approaches in order to evaluate and quantify sequence-dependent dynamics.

% what we did in this paper
\textbf{Key Results.}  In this work, we establish new understanding of the sequence-dependent hybridization/dehybridization kinetics and mechanism within a family of self-complementary pairs of 10-mer DNA oligomers by integrating coarse-grained molecular simulation, machine learning, data-driven inference of long-time kinetic models, and experimental temperature-jump infrared spectroscopy. We conducted more than 1 ms of coarse-grained molecular dynamics simulations and employed deep learning techniques to construct high-resolution Markov state models as predictive and interpretable models of the sequence dependent dynamics. Our results reveal that the specific placement of interrupting G:C pairs within an otherwise repetitive AT sequence can have a profound impact on the kinetic pathways and mechanisms for association and dissociation of the DNA duplex.

% impact and importance of this paper
\textbf{Significance.} Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization, offer quantitative models of their long-time kinetics, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation in burgeoning DNA nanotechnology applications.

% fit for the journal
\textbf{Scope.}  This work fits within the scope of \textit{The Journal of the American Chemical Society} in that it reports new fundamental physical chemical understanding of the molecular pathways and mechanisms of sequence-dependent DNA hybridization/dehybridization pathways. These sequence-dependent mechanisms are of both fundamental interest in biophysical and biochemical processes and of increasing importance in burgeoning DNA nanotechnology applications reliant on quantitative understanding of the sequence-dependent kinetic pathways and rates.

@@

\textbf{\textit{Simulation.}}

@@

$\circ$ \textit{Expert in multi-scale biomolecular simulation with a focus in materials science and biology.}

Jon K. Whitmer \hfill Chemical and Biomolecular Engineering, Notre Dame \\
(574) 631-1417 \hfill \url{jwhitme1@nd.edu}

$\circ$ \textit{Expert in simulations of biomolecular and self-assembling systems.}

@@

$\circ$ \textit{Expert in theory and simulation of coarse-grained biomolecular systems.}

@@

$\circ$ \textit{Expert in simulations of coarse-grained modeling of DNA-protein systems.}

@@

$\circ$ \textit{Expert in simulations of biochemical systems for nanotechnology applications.}

@@

$\circ$ \textit{Expert in molecular dynamics simulations of nucleic acids.}
